---
title: "DSC5101 Group Project 3 Estimating Price Discrimination of Pumpkin"
subtitle: by Dong Gang, Swetha Narayanan, Wang Shenghao, Wang Shuai Max
output:
  html_document:
    highlight: tango
    theme: yeti
  pdf_document: default
  word_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Load Data
The dataset used in this project is stored in **agg_pumpkin_price.csv**. Import the data and remove the invalid entries.

```{r}
raw.data <- read.csv("agg_pumpkin_price.csv")
rate_truck <- 0.37
rate_rail <- 0.03
rate_air <- 4.63
rate_water <- 0.10
lb_to_ton <- 0.0005
km_to_ml <- 0.621371
pumpkin.data <- raw.data[raw.data$Variety!="" & raw.data$Item.Size!="" & raw.data$Origin!="" & raw.data$Package!='each' & raw.data$Package!='bins',]
pumpkin.data <- pumpkin.data[pumpkin.data$Origin!='CANADA' & pumpkin.data$Origin!='MEXICO' & pumpkin.data$Origin!='COSTA RICA',]
pumpkin.data$Variety <- as.factor(pumpkin.data$Variety)
pumpkin.data$Item.Size <- as.factor(pumpkin.data$Item.Size)
pumpkin.data$City.Name <- as.factor(pumpkin.data$City.Name)
pumpkin.data$Origin <- as.factor(pumpkin.data$Origin)
pumpkin.data$avg_price <- (pumpkin.data$Low.Price + pumpkin.data$High.Price) / 2
pumpkin.data$wt_in_ton <- pumpkin.data$Weight * lb_to_ton
pumpkin.data$dist_in_ml <- pumpkin.data$Distance * km_to_ml
pumpkin.data$shipping_cost <- pumpkin.data$dist_in_ml * pumpkin.data$wt_in_ton * rate_truck
pumpkin.data$net_price <- pumpkin.data$avg_price - pumpkin.data$shipping_cost
pumpkin.data <- pumpkin.data[pumpkin.data$net_price > 0,]
```

## Derive Baseline Model with Linear Regression
The following assumptions are made to derive the baseline model.
1. The time effect on pumpkin price is negligible.
2. Shipment of pumpkins is via truck, with a uniform rate of 0.37 US$/(ton.ml).

Hypothesis: Price is same across all the cities
```{r}
price.base.mod <- lm(net_price ~ Variety + Item.Size, data = pumpkin.data)
summary(price.base.mod)

price.base.mod.res <- lm(resid(price.base.mod) ~ City.Name, data = pumpkin.data)
summary(price.base.mod.res)
```
However, it can be observed that the residuals of the net price difference have significant relations with cities.
The initial hypothesis that there is no price difference between different cities fails. 


Effect of cities incomes and CPI on price discrimination.
```{r}
income.data <- read.csv("Income.csv")
income.data <- income.data[income.data$City.Name != "MIAMI",]
income.data$coef <- as.numeric(coef(price.base.mod.res))
income.model <- lm(coef ~ Income, data = income.data)
summary(income.model)

income.cpimodel <- lm(coef ~ CPI, data = income.data)
summary(income.cpimodel)
```


Comparative Study on same origin (Michigan)
```{r}
michigan.data <- pumpkin.data[pumpkin.data$Origin=='MICHIGAN',]
michigan.lm.mod1 <- lm(net_price ~ City.Name + Weight + Variety + Item.Size, data = michigan.data)
summary(michigan.lm.mod1)
```


Second Degree Price Discrimination over pumpkin packages
Hypothesis: all packages share the same unit price
```{r}
#find the Unit Price
pumpkin.data$unitprice <- pumpkin.data$net_price / pumpkin.data$Weight
price.lm.mod3 <- lm(unitprice ~ Variety + City.Name, data = pumpkin.data)
summary(price.lm.mod3)
price.lm.mod3.res <- lm(resid(price.lm.mod3) ~ Package, data = pumpkin.data)
summary(price.lm.mod3.res)
```
The results indicate that the unit price of pumpkins is significantly different across different Packages. 
The hypothsis is rejected.


## Robustness Tests
1. Robustness test with respect to variation of shippping cost
```{r}
pumpkin.data$Date <- as.Date(pumpkin.data$Date, format = "%m/%d/%Y")
pumpkin.data$shipping_cost2 <- pumpkin.data$dist_in_ml * pumpkin.data$wt_in_ton * pumpkin.data$Rate
pumpkin.data$net_price2 <- pumpkin.data$avg_price - pumpkin.data$shipping_cost

price2.lm.mod2 <- lm(net_price2 ~ Variety + Item.Size, data = pumpkin.data)
summary(price2.lm.mod2)
price2.lm.mod2.res <- lm(resid(price2.lm.mod2) ~ City.Name, data = pumpkin.data)
summary(price2.lm.mod2.res)

income.data$coef2 <- as.numeric(coef(price2.lm.mod2.res))
income.model2 <- lm(coef2 ~ Income, data = income.data)
summary(income.model2)
income.cpimodel2 <- lm(coef2 ~ CPI, data = income.data)
summary(income.cpimodel2)

pumpkin.data$unitprice <- pumpkin.data$net_price2 / pumpkin.data$Weight
price2.lm.mod3 <- lm(unitprice ~ Variety + City.Name, data = pumpkin.data)
summary(price2.lm.mod3)
price2.lm.mod3.res <- lm(resid(price2.lm.mod3) ~ Package, data = pumpkin.data)
summary(price2.lm.mod3.res)

```


2. Robustness test with respect to time effect
```{r}
library(plm)
library(Formula)
pumpkin.data$Date <- as.Date(pumpkin.data$Date, format = "%m/%d/%Y")
price.lm.mod2 <- plm(net_price ~ Variety + Item.Size, data = pumpkin.data, index = c("Date"), model = "within")
summary(price.lm.mod2)

price.lm.mod2.res <- plm(resid(price.lm.mod2) ~ City.Name, data = pumpkin.data, index = c("Date"), model = "within")
summary(price.lm.mod2.res)
```


## Demand Elasticity Analysis
```{r}
#Demand Elasticity
nrow(pumpkin.data)

lastyear.data <- pumpkin.data[which(pumpkin.data$Year == "2016" & pumpkin.data$City.Name == "SAN FRANCISCO" ) ,]
thisyear.data <- pumpkin.data[which(pumpkin.data$Year == "2017" & pumpkin.data$City.Name == "SAN FRANCISCO" ) ,]
lastyear.data <- pumpkin.data[which(pumpkin.data$Year == "2016" & pumpkin.data$City.Name == "COLUMBIA" ) ,]
thisyear.data <- pumpkin.data[which(pumpkin.data$Year == "2017" & pumpkin.data$City.Name == "COLUMBIA" ) ,]


Q1 <- sum(lastyear.data$Q)
Price1 <- sum(lastyear.data$avg_price)/nrow(lastyear.data)
Q2 <- sum(thisyear.data$Q)
Price2 <- sum(thisyear.data$avg_price)/nrow(thisyear.data)
e1 = ((Q2 - Q1)/Q1)/((Price2 - Price1)/Price1)
e1
#SF inelastic - 0.88
#Columbia - Elastic - 4.856638
#Boston - Elastic - 3.827553
#New york - 3.392982
#LA - Inelastic - 0.4743102

plot(lastyear.data$Month,lastyear.data$avg_price)
unique(lastyear.data$Month)

lastyear.september.data <- lastyear.data[lastyear.data$Month == "9",]
lastyear.october.data <- lastyear.data [lastyear.data$Month  == "10",]
lastyear.november.data <- lastyear.data [lastyear.data$Month  == "11",]
lastyear.december.data <- lastyear.data [lastyear.data$Month  == "12",]
```

***[THE END]**